\bf Model Visualization Techniques for a Social Network Model

### 2018-09-12 paper-child.Rnw
just do it.

In addition to assessing model fit, we also want to learn about model behavior so that we can choose better fitting models in the future. We examine the behavior of SAOMs by visualizing collections of models. We build collections of models in four ways. First, we explore the space of all possible models to determine the most significant model parameters.  Second, we vary model settings to uncover relationships between parameters and their effects on the fitted model. Next, we fit the same model to the same data many times to assess the behavior of the underlying simulations. Finally, we fit the same model to different data to learn about fitted parameter values under different conditions. We chose these four collections because they each explore something different about SAOM behavior: the choice of parameters in the objective function, the effects of underlying simulations, and the resulting fitted values under different circumstances. Using the tools in our visualization toolbox, we %The first considers the many parameters to include in a SAOM, while the second results in different parameter estimates, so we can see how parameter estimates vary. The third collection shows how those many possible parameters affect the values of parameters in the fitted models. Finally, we explore how the same model behaves under two completely different circumstances.

@@

\noindent \textbf{Exploring the space of all possible models:} The \texttt{RSiena} manual contains over eighty effects to include in a SAOM's objective function. In order to select parameters to include in the models for our example data, we searched through the possible effects available to model given the data structure to find \textit{significant} effects. We tested for significance using the Wald-type tests built into \texttt{RSiena} for one-at-a-time effects testing. We start with the outdegree and reciprocity measures as the foundation of the models we fit, then add one effect, fit the model, test the additional effect for significance, and repeat for all possible parameters to add to the model. We performed this procedure, which we used to pick M2 and M3, for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in Figure~\ref{fig:sigtesting} in the Appendix. %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

@@

\noindent \textbf{Varying model settings:} The next method of viewing collections of models is with varying model settings. We have varied model settings already by choosing three different models M1, M2, M3 to fit to our example data sets. In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examined the parameter densities. Here, we simulate from these three models. For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model. We simulated 1,000 observations from each of the three models in this way. From these simulations, we created visualizations that represent an average network from models M1, M2, and M3 using the same method described in Section~\ref{sec:minds}. These averages are shown in Figure~\ref{fig:averagenet} in the Appendix. We see that all models are failing to capture the structure of the second wave.

>>>>

Now that we have assessed model fit, we next want to learn about model behavior in order to choose better fitting models in the future. We examine the behavior of SAOMs by visualizing collections of models. We build collections of models in four ways. First, we explore the space of all possible models to determine the most significant model parameters.  Second, we vary model settings to uncover relationships between parameters and their effects on the fitted model. Next, we fit the same model to the same data many times to assess the behavior of the underlying simulations. Finally, we fit the same model to different data to learn about fitted parameter values under different conditions. We chose these four collections because they each explore something different about SAOM behavior: the choice of parameters in the objective function, the effects of underlying simulations, and the resulting fitted values under different circumstances. Using the tools in our visualization toolbox, we %The first considers the many parameters to include in a SAOM, while the second results in different parameter estimates, so we can see how parameter estimates vary. The third collection shows how those many possible parameters affect the values of parameters in the fitted models. Finally, we explore how the same model behaves under two completely different circumstances.

@@

\noindent \textbf{Exploring the space of all possible models:} The \texttt{RSiena} manual contains over eighty effects to include in the SAOM objective function. To select objective function parameters for fitting models to our example data, we searched through all possible effects available to find \textit{significant} effects. The \texttt{RSiena} package includes the function \texttt{Wald.Rsiena()} for Wald-type tests of parameter significance. We start by including the outdegree and reciprocity parameters in the model then add one effect to the objective function, fit the model to the example data, test the added effect for significance, and repeat for all possible model parameters. We performed this procedure, which we used to pick M2 and M3, for both the small friendship and the senate collaboration data. A visualization of the significant effects for the two example data sets are shown in Figure~\ref{fig:sigtesting} in the Appendix. Using this one-at-a-time testing method results in many possible significant parameters of different type and magnitude, demonstrating the flexibility of these models.  %We see in both the friendship data and the senate data results that most of the significant effects have absolute value less than ten. In addition, the $p$-values for the effects from the friendship data are more spread out than the $p$-values for the senate data, which are concentrated at about 0.02 or less. This may suggest that larger data sets tend to result generally in smaller $p$-values, just like a larger sample size results in smaller $p$-values in a $t$-test.

@@

\noindent \textbf{Varying model settings:} We now view collections of models by varying model settings. We have done this already by choosing three different models M1, M2, M3 to fit to our example data sets. To explore the behavior of these three models, we simulate from them. Using the parameter values given in Table~\ref{tab:meanests} for the friendship data, we simulated 1,000 observations from each of the three models and created expected network visualizations from models M1, M2, and M3 using the same method as in Section~\ref{sec:minds}. These averages are shown in Figure~\ref{fig:averagenet} in the Appendix. We see that all models are failing to capture the structure of the second wave. %In Section~\ref{sec:samemodsamedat}, we fit these three models to the data 1,000 times, and examine the parameter densities. Here, . For simulation, we used the mean parameter values from the 1,000 fits of the corresponding model.
